\vspace{1.0cm} \LARGE \bf Supporting Text: \\ Revealing the competitive dynamics of influenza viruses through evolutionary cartography

### 2013-01-22 flux_si.tex
Fix math notation problem (but this estimator still may not be correct).

Owing to this relationship, Pybus et al.\ \cite{Pybus12} estimate $D$ from phylogenies in which internal nodes have inferred 2D locations via the formula
\begin{equation} \label{pybusest}
	\hat{D} = \frac{1}{n} \sum^n_{i=1} \frac{d_i^2}{4t_i},
\end{equation}
in which $d_i$ represents the displacement of a branch and $t_i$ represents the length of a branch.
However, short branches along the phylogeny may cause $d_i^2/t_i$ to have high variance and lead to uncertainty in the resulting estimates of $D$.
Consequently, we estimate $D$ by looking at the relationship between total squared displacement along branches of a phylogeny and total time along branches of a phylogeny
\begin{eqnarray}
	d_\mathrm{total} = \frac{1}{n} \sum^n_{i=1} d_i^2 \nonumber \\
	t_\mathrm{total} = \frac{1}{n} \sum^n_{i=1} t_i^2 \nonumber \\
	\hat{D} = \frac{ d_\mathrm{total} }{ 4 t_\mathrm{total} }.
\end{eqnarray}
This estimator showed less variance than the estimator in equation \ref{pybusest} when analyzing tree output from the BMDS model and was thus preferred.
When estimating $D$ specific to trunk branches or side branches, we calculate $d_\mathrm{total}$ and $t_\mathrm{total}$ only from the subset of trunk branches or side branches as appropriate.

>>>>

Owing to this relationship, Pybus et al.\ \cite{Pybus12} estimate $D$ from phylogenies in which internal nodes have inferred 2D locations via the formula
\begin{equation} \label{pybusest}
	\hat{D} = \frac{1}{n} \sum^n_{i=1} \frac{d_i^2}{4t_i},
\end{equation}
in which $d_i$ represents the displacement of a branch and $t_i$ represents the length of a branch.
However, short branches along the phylogeny may cause $d_i^2/t_i$ to have high variance and lead to uncertainty in the resulting estimates of $D$.
Consequently, we estimate $D$ by looking at the relationship between total squared displacement along branches of a phylogeny and total time along branches of a phylogeny
\begin{eqnarray}
	d^2_\mathrm{total} = \sum^n_{i=1} d^2_i \nonumber \\
	t_\mathrm{total} = \sum^n_{i=1} t_i^2 \nonumber \\
	\hat{D} = \frac{ d^2_\mathrm{total} }{ 4 t_\mathrm{total} }.
\end{eqnarray}
This estimator showed less variance than the estimator in equation \ref{pybusest} when analyzing tree output from the BMDS model and was thus preferred.
When estimating $D$ specific to trunk branches or side branches, we calculate $d^2_\mathrm{total}$ and $t_\mathrm{total}$ only from the subset of trunk branches or side branches as appropriate.
